---
title: "CZU_grouping"
author: "Bryant Willoughby"
date: "2023-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(readr)
library(ggplot2)
source("C:/Users/Bryant Willoughby/Documents/DRUMS/code/week 10/boxplot_function.R")
  #boxplot function 
source("C:/Users/Bryant Willoughby/Documents/DRUMS/code/week 10/regression_transformation.R")
  #linear regression transformation function 
firedate <- as.Date("2020-08-16")
  #day of fire ignition 
```



Read in final data
```{r}
data <- read_csv("C:/Users/Bryant Willoughby/Documents/DRUMS/raw(CZU)/Final Data Sets/prefire_czu_finaldata.csv")

```

# Transform variables taken across different days with linear regression
- Note: 8/16/2020 - lightning strike ignition
  

Explore LST variable 
- notice: 08/08 measured twice 
```{r}
LST <- data %>%
  select(starts_with("LST"))

names(LST)

var_boxplot(LST)
```

combine two measurements taken on 8/8 by averaging and replacing 
```{r}
LST <- LST %>%
  mutate("LST_2020.08.08.00.00.00" = rowMeans(select(., c("LST_2020.08.08.03.00.00", 
                                                          "LST_2020.08.08.19.00.00")), 
                                              na.rm = TRUE)) %>%
  select(-c("LST_2020.08.08.03.00.00", "LST_2020.08.08.19.00.00"))

```

```{r}
var_boxplot(LST)
```


```{r}
result <- var_tran(
  dates = as.Date(c("2020-08-04", "2020-08-08", "2020-08-12", "2020-08-15", "2020-08-16")), 
  var = LST
)

# Store the matrix X in LST_sum
LST_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(LST_sum) <- c("LST_int", "LST_slope")
```


ET variable
```{r}
ET <- data %>%
  select(starts_with("ET"))

names(ET)

var_boxplot(ET)
```


```{r}
result <- var_tran(dates = as.Date(c("2020-08-04", "2020-08-08", "2020-08-15", "2020-08-16")), 
                   var = ET
)

# Store the matrix X in LST_sum
ET_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(ET_sum) <- c("ET_int", "ET_slope")

```

ESI variable 
```{r}
ESI <- data %>%
  select(starts_with("ESI"))

names(ESI)

var_boxplot(ESI)
```

```{r}
result <- var_tran(dates = as.Date(c("2020-08-04", "2020-08-08", "2020-08-15", "2020-08-16")), 
                   var = ESI
)

# Store the matrix X in LST_sum
ESI_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(ESI_sum) <- c("ESI_int", "ESI_slope")
```

WUE
```{r}
WUE <- data %>%
  select(starts_with("WUE"))

names(WUE)

var_boxplot(WUE)
```

```{r}
result <- var_tran(dates = as.Date(c("2020-08-04", "2020-08-08", "2020-08-15", "2020-08-16")), 
                   var = WUE
)

# Store the matrix X in LST_sum
WUE_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(WUE_sum) <- c("WUE_int", "WUE_slope")
```

LAI
```{r}
LAI <- data %>%
  select(starts_with("LAI"))

names(LAI)

var_boxplot(LAI)
```

```{r}
result <- var_tran(dates = as.Date(c("2020-08-05", "2020-08-13")), 
                   var = LAI
)

# Store the matrix X in LST_sum
LAI_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(LAI_sum) <- c("LAI_int", "LAI_slope")
```

NDVI
```{r}
NDVI <- data %>%
  select(starts_with("NDVI"))

names(NDVI)

var_boxplot(NDVI)
```


```{r}
result <- var_tran(dates = as.Date(c("2020-08-03", "2020-08-08", "2020-08-10",
                                     "2020-08-13")), 
                   var = NDVI
)

# Store the matrix X in LST_sum
NDVI_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(NDVI_sum) <- c("NDVI_int", "NDVI_slope")
```


AM_soil_moisture
```{r}
AM_sm <- data %>%
  select(starts_with("AM"))

names(AM_sm)

var_boxplot(AM_sm)
```

```{r}
result <- var_tran(dates = as.Date(c("2020-08-04", "2020-08-05", "2020-08-07",
                   "2020-08-10", "2020-08-12", "2020-08-13", 
                   "2020-08-15")), 
                   var = AM_sm
)

# Store the matrix X in LST_sum
AM_sm_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(AM_sm_sum) <- c("AM_sm_int", "AM_sm_slope")
```


PM soil moisture 
```{r}
PM_sm <- data %>%
  select(starts_with("soil"))

names(PM_sm)

var_boxplot(PM_sm)
```


```{r}
result <- var_tran(dates = as.Date(c("2020-08-03", "2020-08-06", "2020-08-08",
                   "2020-08-09", "2020-08-11", "2020-08-14", 
                   "2020-08-16", "2020-08-17")), 
                   var = PM_sm
)

# Store the matrix X in LST_sum
PM_sm_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(PM_sm_sum) <- c("PM_sm_int", "PM_sm_slope")
```


Dewpoint temp (tdmean)
```{r}
tdmean <- data %>%
  select(starts_with("tdmean"))

names(tdmean)

var_boxplot(tdmean)
```

```{r}
result <- var_tran(dates = as.Date(c("2020-08-02", "2020-08-03", "2020-08-04",
                   "2020-08-05", "2020-08-06", "2020-08-07", 
                   "2020-08-08", "2020-08-09", "2020-08-10", 
                   "2020-08-11", "2020-08-12", "2020-08-13", 
                   "2020-08-14", "2020-08-15", "2020-08-16")), 
                   var = tdmean
)

# Store the matrix X in LST_sum
tdmean_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(tdmean_sum) <- c("tdmean_int", "tdmean_slope")
```

Air temp (tmean)
```{r}
tmean <- data %>%
  select(starts_with("tmean"))

names(tmean)

var_boxplot(tmean)
```

```{r}
result <- var_tran(dates = as.Date(c("2020-08-02", "2020-08-03", "2020-08-04",
                   "2020-08-05", "2020-08-06", "2020-08-07", 
                   "2020-08-08", "2020-08-09", "2020-08-10", 
                   "2020-08-11", "2020-08-12", "2020-08-13", 
                   "2020-08-14", "2020-08-15", "2020-08-16")), 
                   var = tmean
)

# Store the matrix X in LST_sum
tmean_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(tmean_sum) <- c("tmean_int", "tmean_slope")
```

Vapor Pressure Max (vpdmax)
```{r}
vpdmax <- data %>%
  select(starts_with("vpdmax"))

names(vpdmax)

var_boxplot(vpdmax)
```

```{r}
result <- var_tran(dates = as.Date(c("2020-08-02", "2020-08-03", "2020-08-04",
                   "2020-08-05", "2020-08-06", "2020-08-07", 
                   "2020-08-08", "2020-08-09", "2020-08-10", 
                   "2020-08-11", "2020-08-12", "2020-08-13", 
                   "2020-08-14", "2020-08-15", "2020-08-16")), 
                   var = vpdmax
)

# Store the matrix X in LST_sum
vpdmax_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(vpdmax_sum) <- c("vpdmax_int", "vpdmax_slope")

```


Vapor Pressure Min (vpdmin)
```{r}
vpdmin <- data %>%
  select(starts_with("vpdmin"))

names(vpdmin)

var_boxplot(vpdmin)
```

```{r}
result <- var_tran(dates = as.Date(c("2020-08-02", "2020-08-03", "2020-08-04",
                   "2020-08-05", "2020-08-06", "2020-08-07", 
                   "2020-08-08", "2020-08-09", "2020-08-10", 
                   "2020-08-11", "2020-08-12", "2020-08-13", 
                   "2020-08-14", "2020-08-15", "2020-08-16")), 
                   var = vpdmin
)

# Store the matrix X in LST_sum
vpdmin_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(vpdmin_sum) <- c("vpdmin_int", "vpdmin_slope")

```

Create df w/ slope & intercept term
```{r}
newdata <- data %>%
  select("Lat", "Lon", "dnbr6", "dnbr", 
         "aspect", "slope", "elev", "LC")

fulldata <- cbind(newdata, LST_sum, ET_sum, ESI_sum, 
                  WUE_sum, LAI_sum, NDVI_sum, AM_sm_sum, 
                  PM_sm_sum, tdmean_sum, tmean_sum, vpdmax_sum, 
                  vpdmin_sum)

```


Write to CSV
```{r}
write.csv(fulldata, "C:/Users/Bryant Willoughby/Documents/DRUMS/raw(CZU)/Final Data Sets/prefire_czu_grouped_finaldata.csv", row.names = F)
```


```{r}
finalgrouped <- read_csv("C:/Users/Bryant Willoughby/Documents/DRUMS/raw(CZU)/Final Data Sets/prefire_czu_grouped_finaldata.csv")

```



