---
title: "Kincade_Grouping"
author: "Bryant Willoughby"
date: "2023-08-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(readr)
library(ggplot2)
source("C:/Users/Bryant Willoughby/Documents/DRUMS/code/week 10/boxplot_function.R")
  #boxplot function 
source("C:/Users/Bryant Willoughby/Documents/DRUMS/code/week 10/regression_transformation.R")
  #linear regression transformation function 
firedate <- as.Date("2019-10-23")
  #day of fire ignition 
```

Read in final data
```{r}
data <- read_csv("C:/Users/Bryant Willoughby/Documents/DRUMS/raw(Kincade)/Final Data Sets/prefire_kincade_finaldata.csv")

```


# Transform variables taken across different days with linear regression
- Note: 10/23/2019  - lightning strike ignition
  

Explore LST variable 
- notice: 08/08 measured twice 
```{r}
LST <- data %>%
  select(starts_with("LST"))

names(LST)

var_boxplot(LST)
```


```{r}
result <- var_tran(
  dates = as.Date(c("2019-10-15", "2019-10-22")), 
  var = LST
)

# Store the matrix X in LST_sum
LST_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(LST_sum) <- c("LST_int", "LST_slope")
```


ET variable
```{r}
ET <- data %>%
  select(starts_with("ET"))

names(ET)

var_boxplot(ET)
```


```{r}
result <- var_tran(dates = as.Date(c("2019-10-15", "2019-10-22")), 
                   var = ET
)

# Store the matrix X in LST_sum
ET_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(ET_sum) <- c("ET_int", "ET_slope")

```


ESI variable 
```{r}
ESI <- data %>%
  select(starts_with("ESI"))

names(ESI)

var_boxplot(ESI)
```



```{r}
result <- var_tran(dates = as.Date(c("2019-10-15", "2019-10-22")), 
                   var = ESI
)

# Store the matrix X in LST_sum
ESI_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(ESI_sum) <- c("ESI_int", "ESI_slope")
```


WUE
```{r}
WUE <- data %>%
  select(starts_with("WUE"))

names(WUE)

var_boxplot(WUE)
```

```{r}
result <- var_tran(dates = as.Date(c("2019-10-15", "2019-10-22")), 
                   var = WUE
)

# Store the matrix X in LST_sum
WUE_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(WUE_sum) <- c("WUE_int", "WUE_slope")
```


LAI
```{r}
LAI <- data %>%
  select(starts_with("LAI"))

names(LAI)

var_boxplot(LAI)
```


```{r}
result <- var_tran(dates = as.Date(c("2019-10-09", "2019-10-13", "2019-10-17", "2019-10-21")), 
                   var = LAI
)

# Store the matrix X in LST_sum
LAI_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(LAI_sum) <- c("LAI_int", "LAI_slope")
```


NDVI
```{r}
NDVI <- data %>%
  select(starts_with("NDVI"))

names(NDVI)

var_boxplot(NDVI)
```


```{r}
result <- var_tran(dates = as.Date(c("2019-10-11", "2019-10-13", "2019-10-16", "2019-10-18", "2019-10-21", 
                                     "2019-10-23")), 
                   var = NDVI
)

# Store the matrix X in LST_sum
NDVI_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(NDVI_sum) <- c("NDVI_int", "NDVI_slope")
```

AM_soil_moisture
```{r}
AM_sm <- data %>%
  select(starts_with("AM"))

names(AM_sm)

var_boxplot(AM_sm)
```


```{r}
result <- var_tran(dates = as.Date(c("2019-10-11", "2019-10-13", "2019-10-14", "2019-10-16", "2019-10-19", 
                                     "2019-10-21", "2019-10-22", "2019-10-24")), 
                   var = AM_sm
)

# Store the matrix X in LST_sum
AM_sm_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(AM_sm_sum) <- c("AM_sm_int", "AM_sm_slope")
```


PM soil moisture 
```{r}
PM_sm <- data %>%
  select(starts_with("soil"))

names(PM_sm)

var_boxplot(PM_sm)
```


```{r}
result <- var_tran(dates = as.Date(c("2019-10-12", "2019-10-15", "2019-10-17", "2019-10-18", "2019-10-20", 
                                     "2019-10-23")), 
                   var = PM_sm
)

# Store the matrix X in LST_sum
PM_sm_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(PM_sm_sum) <- c("PM_sm_int", "PM_sm_slope")
```


Dewpoint temp (tdmean)
```{r}
tdmean <- data %>%
  select(starts_with("tdmean"))

names(tdmean)

var_boxplot(tdmean)
```


```{r}
result <- var_tran(dates = as.Date(c("2019-10-11", "2019-10-12", "2019-10-13", "2019-10-14", "2019-10-15", 
                                     "2019-10-16", "2019-10-17", "2019-10-18", "2019-10-19", "2019-10-20", 
                                     "2019-10-21", "2019-10-22", "2019-10-23", "2019-10-24")), 
                   var = tdmean
)

# Store the matrix X in LST_sum
tdmean_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(tdmean_sum) <- c("tdmean_int", "tdmean_slope")
```


Air temp (tmean)
```{r}
tmean <- data %>%
  select(starts_with("tmean"))

names(tmean)

var_boxplot(tmean)
```


```{r}
result <- var_tran(dates = as.Date(c("2019-10-11", "2019-10-12", "2019-10-13", "2019-10-14", "2019-10-15", 
                                     "2019-10-16", "2019-10-17", "2019-10-18", "2019-10-19", "2019-10-20", 
                                     "2019-10-21", "2019-10-22", "2019-10-23", "2019-10-24")), 
                   var = tmean
)

# Store the matrix X in LST_sum
tmean_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(tmean_sum) <- c("tmean_int", "tmean_slope")
```


Vapor Pressure Max (vpdmax)
```{r}
vpdmax <- data %>%
  select(starts_with("vpdmax"))

names(vpdmax)

var_boxplot(vpdmax)
```


```{r}
result <- var_tran(dates = as.Date(c("2019-10-11", "2019-10-12", "2019-10-13", "2019-10-14", "2019-10-15", 
                                     "2019-10-16", "2019-10-17", "2019-10-18", "2019-10-19", "2019-10-20", 
                                     "2019-10-21", "2019-10-22", "2019-10-23", "2019-10-24")), 
                   var = vpdmax
)

# Store the matrix X in LST_sum
vpdmax_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(vpdmax_sum) <- c("vpdmax_int", "vpdmax_slope")

```


Vapor Pressure Min (vpdmin)
```{r}
vpdmin <- data %>%
  select(starts_with("vpdmin"))

names(vpdmin)

var_boxplot(vpdmin)
```


```{r}
result <- var_tran(dates = as.Date(c("2019-10-11", "2019-10-12", "2019-10-13", "2019-10-14", "2019-10-15", 
                                     "2019-10-16", "2019-10-17", "2019-10-18", "2019-10-19", "2019-10-20", 
                                     "2019-10-21", "2019-10-22", "2019-10-23", "2019-10-24")), 
                   var = vpdmin
)

# Store the matrix X in LST_sum
vpdmin_sum <- data.frame(result)

# Name the columns of LST_sum
colnames(vpdmin_sum) <- c("vpdmin_int", "vpdmin_slope")

```


Create df w/ slope & intercept term
```{r}
newdata <- data %>%
  select("Lat", "Lon", "dnbr6", "dnbr", 
         "aspect", "slope", "elev", "LC")

fulldata <- cbind(newdata, LST_sum, ET_sum, ESI_sum, 
                  WUE_sum, LAI_sum, NDVI_sum, AM_sm_sum, 
                  PM_sm_sum, tdmean_sum, tmean_sum, vpdmax_sum, 
                  vpdmin_sum)

```

Write to CSV
```{r}
write.csv(fulldata, "C:/Users/Bryant Willoughby/Documents/DRUMS/raw(Kincade)/Final Data Sets/prefire_kincade_grouped_finaldata.csv", row.names = F)
```






