---
title: "KNP_Independent"
author: "Bryant Willoughby"
date: "2023-08-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(SuperLearner)
library(GpGp)
library(ggplot2)
library(caret)
```

Read in fulldata
```{r}
fulldata <- read_csv("C:/Users/Bryant Willoughby/Documents/DRUMS/raw(KNP)/Final data sets/prefire_knp_blr_grouped_standard_finaldata.csv")
```

Explore dnbr6 variable
- note: dnbr6 = 4 is the highest burn severity classification
```{r}
#view unique levels of dnbr6
unique_levels <- unique(fulldata$dnbr6)
unique_levels #1,2,3,4,5

#view range of variable 
range_summary <- fulldata %>%
  group_by(dnbr6) %>%
  summarize(min_dnbr = min(dnbr), max_dnbr = max(dnbr))

#move in ascending order
updated_summary <- range_summary %>%
  mutate(order_var = ifelse(dnbr6 == 5, -Inf, 0)) %>%
  arrange(order_var) %>%
  select(-order_var)

updated_summary

```

Read in training/test data 
```{r}
train <- read_csv("C:/Users/Bryant Willoughby/Documents/DRUMS/raw(KNP)/intermediate/KNP_traindata.csv")

test <- read_csv("C:/Users/Bryant Willoughby/Documents/DRUMS/raw(KNP)/intermediate/KNP_testdata.csv")
```


Linear Regression (Independent covariance) Results 
```{r}
#print selected variables 
unwanted_columns <- c("LC", "Lat", "Lon", "dnbr")
selected_variables <- setdiff(names(train), unwanted_columns)
  #extract elements unique to train data names

#fit the final model using selected variables 
final_model <- lm(dnbr ~ ., data = train[, c("dnbr", selected_variables)])

#anova(final_model)
summary(final_model) #about 20.33% of variance explained by model 

```

Use fitted coefficients to make predictions on test data 
```{r}
#make dnbr preds 
predicted_values <- predict(final_model, newdata = test[, !(names(test) %in% c("LC", "Lat", "Lon", "dnbr"))])

preds <- as.vector(predicted_values)
```

# RMSE calculation
```{r}
source("C:/Users/Bryant Willoughby/Documents/DRUMS/code/week a/RMSE_function.R")
actual <- test$dnbr
predicted <- preds

val_RMSE <- RMSE(actual, predicted)

val_RMSE <- round(val_RMSE,0)
val_RMSE
```

# Create confusion matrix
Use defined ordinal severity levels
- enhanced regrowth (high): (-500,-251)
- enhanced regrowth (lo): (-250,-101)
- Unburned: (-100,99)
- Low: (100,269)
- Moderate-low: (270,439)
- Moderate-high: (440,659)
- High (660,1300)
note: dnbr <-550 or >1350 are likely anomalies not related to real land cover differences

```{r}
range(fulldata$dnbr)
  #not: don't need to worry about anomalies
```

Create summary table for confusion matrix computation
```{r}
data <- test[2:4] %>%
  mutate("dnbr_class" = if_else(dnbr >= 660, 7, 
                                if_else(dnbr >= 440, 6, 
                                if_else(dnbr >= 270, 5,
                                if_else(dnbr >= 100, 4,
                                if_else(dnbr >= -100, 3, 
                                if_else(dnbr >= -250, 2, 1)))))), 
         "dnbr_preds" = preds, 
         "dnbr_preds_class" = if_else(preds >= 660, 7, 
                                if_else(preds >= 440, 6, 
                                if_else(preds >= 270, 5,
                                if_else(preds >= 100, 4,
                                if_else(preds >= -100, 3, 
                                if_else(preds >= -250, 2, 1)))))))

```


```{r}
#Creates vectors having data points
expected_val <- factor(data$dnbr_class)
predicted_val <- factor(data$dnbr_preds_class)

#Creating confusion matrix
confusion_mat <- confusionMatrix(data=predicted_val, reference = expected_val)

confusion_mat
  #note: Accuracy = number of correct predictions / total number of predictions
```

Classification accuracy (CA)
```{r}
CA <- confusion_mat$overall[[1]]
CA <- round(CA, 3)
CA <- CA*100
CA
```

Classification accuracy high (CA-high)
```{r}
CA_high <- confusion_mat$table[7,7] / sum(confusion_mat$table[,7])
CA_high <- round(CA_high, 4)
CA_high <- CA_high*100
CA_high
```








