---
title: "CZU_EDA"
author: "Bryant Willoughby"
date: "2023-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
```

Read in Data 

```{r}
data <- read_csv("C:/Users/Bryant Willoughby/Documents/DRUMS/raw(CZU)/Final Data Sets/prefire_czu_fulldata.csv")

```
Check to see if there are any completely missing rows (records)
- there are none
```{r}
head(data)

#keeping rows that have complete information in all columns (except the first four) or have at least one non-missing value in any of the columns (excluding the first four columns).
cleaned <- data[complete.cases(data[,-(1:4)]) | rowSums(!is.na(data[,-(1:4)])) > 0, ]

# Print the updated data frame
head(cleaned)

nrow(data) - nrow(cleaned) #0
```

```{r}
#find num of missing vals
missing_counts <- colSums(is.na(cleaned))

#create df w/ col names and missing values 
missing_data <- data.frame(
  variable = names(missing_counts), 
  missing_values = missing_counts
)

print(missing_data)

```

Map values with more than 5% missing values 
```{r}
# Initialize an empty vector to store the names that satisfy the condition
selected_labels <- c()

# Loop through the list and add the names that meet the condition to the vector
for (label in names(missing_counts)) {
  value <- missing_counts[[label]]
  if (value > 4430) {
    selected_labels <- c(selected_labels, label)
  }
}

# Print the selected labels
print(selected_labels) #26
```

```{r}
map_data <- data %>%
  select(x, y, all_of(selected_labels))

```

```{r}
source("C:/Users/Bryant Willoughby/Documents/DRUMS/code/week 1/maps_function.R")

for (col_name in names(map_data)[3:ncol(map_data)]){
  #note: col_name is a character string of var names
  col <- map_data[[col_name]] #values of each variable 
  
  p <- map.heatmap(map_data$x, map_data$y, col, legendTitle = col_name, 
                   mainTitle = paste(col_name, "across region"))
  print(p)
}

```

Remove variables with >50% (44290 obs) missing data 
```{r}
# Initialize an empty vector to store the names that satisfy the condition
selected_labels <- c()

# Loop through the list and add the names that meet the condition to the vector
for (label in names(missing_counts)) {
  value <- missing_counts[[label]]
  if (value > 44290) {
    selected_labels <- c(selected_labels, label)
  }
}

# Print the selected labels
print(selected_labels)
```
```{r}
map_data <- data %>%
  select(x, y, all_of(selected_labels))

```

```{r}
source("C:/Users/Bryant Willoughby/Documents/DRUMS/code/week 1/maps_function.R")

for (col_name in names(map_data)[3:ncol(map_data)]){
  #note: col_name is a character string of var names
  col <- map_data[[col_name]] #values of each variable 
  
  p <- map.heatmap(map_data$x, map_data$y, col, legendTitle = col_name, 
                   mainTitle = paste(col_name, "across region"))
  print(p)
}

```


```{r}
ncol(cleaned) #119

map_data <- data %>%
  select(x, y, all_of(selected_labels))

#remove variables with 'significant' missing data 
cleaned <- cleaned %>%
  select(-all_of(selected_labels))

ncol(cleaned) #111

```

Check remaining variables with missing values 
```{r}
#find num of missing vals
missing_counts <- colSums(is.na(cleaned))

#create df w/ col names and missing values 
missing_data <- data.frame(
  variable = names(missing_counts), 
  missing_values = missing_counts
)

print(missing_data)

```

```{r}
# Initialize an empty vector to store the names that satisfy the condition
selected_labels <- c()

# Loop through the list and add the names that meet the condition to the vector
for (label in names(missing_counts)) {
  value <- missing_counts[[label]]
  if (value > 4430) {
    selected_labels <- c(selected_labels, label)
  }
}

# Print the selected labels
print(selected_labels) #18
```

```{r}
map_data <- data %>%
  select(x, y, all_of(selected_labels))

```

```{r}
source("C:/Users/Bryant Willoughby/Documents/DRUMS/code/week 1/maps_function.R")

for (col_name in names(map_data)[3:ncol(map_data)]){
  #note: col_name is a character string of var names
  col <- map_data[[col_name]] #values of each variable 
  
  p <- map.heatmap(map_data$x, map_data$y, col, legendTitle = col_name, 
                   mainTitle = paste(col_name, "across region"))
  print(p)
}

```

Removing variables with missing values in same general vicinity 
- most ECOSTRESS vars on 8/12 
```{r}
ncol(cleaned) #111

cleaned <- cleaned %>%
  select(-c("ET_2020.08.12.01.00.00", "ESI_2020.08.12.01.00.00", "WUE_2020.08.12.01.00.00", 
            "LAI_2020.08.09.00.00.00"))

ncol(cleaned) #107
```

Final maps with remaining variables we will impute with KNN
```{r}
#find num of missing vals
missing_counts <- colSums(is.na(cleaned))

#create df w/ col names and missing values 
missing_data <- data.frame(
  variable = names(missing_counts), 
  missing_values = missing_counts
)
```

```{r}
# Initialize an empty vector to store the names that satisfy the condition
selected_labels <- c()

# Loop through the list and add the names that meet the condition to the vector
for (label in names(missing_counts)) {
  value <- missing_counts[[label]]
  if (value > 4430) {
    selected_labels <- c(selected_labels, label)
  }
}

# Print the selected labels
print(selected_labels) #15
```

```{r}
map_data <- data %>%
  select(x, y, all_of(selected_labels))

```

- less missing values comparatively with other variables already removed
- those with more missing values are spread out along region; more suitable for KNN
```{r}
source("C:/Users/Bryant Willoughby/Documents/DRUMS/code/week 1/maps_function.R")

for (col_name in names(map_data)[3:ncol(map_data)]){
  #note: col_name is a character string of var names
  col <- map_data[[col_name]] #values of each variable 
  
  p <- map.heatmap(map_data$x, map_data$y, col, legendTitle = col_name, 
                   mainTitle = paste(col_name, "across region"))
  print(p)
}

```


Perform KNN on dataset
```{r}
source("C:/Users/Bryant Willoughby/Documents/DRUMS/code/week 1/KNN_imputation_function.R")

for (i in 5:ncol(cleaned)){
  cleaned[[i]] <- knn(cleaned$x, cleaned$y, cleaned[[i]])
}

```

confirm there are no more missing values 
```{r}
#find num of missing vals
missing_counts <- colSums(is.na(cleaned))

#create df w/ col names and missing values 
missing_data <- data.frame(
  variable = names(missing_counts), 
  missing_values = missing_counts
)

print(missing_data)

anyNA(cleaned)

```

Rename x/y variables as long/lat respectively
Change LC_2020.01.02.00.00.00 to LC 
```{r}
cleaned <- cleaned %>%
  select(y, x, everything()) %>%
  rename("Lat" = y, "Lon" = x, "LC" = LC_2020.01.02.00.00.00)

```


Write this cleaned data set to a csv file to work with
```{r}
#write.csv(cleaned, file = "C:/Users/Bryant Willoughby/Documents/DRUMS/raw(CZU)/Final Data Sets/prefire_czu_finaldata.csv", row.names = F)
          
```








