---
title: "Kincade_EDA"
author: "Bryant Willoughby"
date: "2023-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
```


Read in Data 

```{r}
data <- read_csv("C:/Users/Bryant Willoughby/Documents/DRUMS/raw(Kincade)/Final Data Sets/prefire_kincade_fulldata.csv")

```

Check to see if there are any completely missing rows (records)
- there are none
```{r}
head(data)

#keeping rows that have complete information in all columns (except the first four) or have at least one non-missing value in any of the columns (excluding the first four columns).
cleaned <- data[complete.cases(data[,-(1:4)]) | rowSums(!is.na(data[,-(1:4)])) > 0, ]

# Print the updated data frame
head(cleaned)

nrow(data) - nrow(cleaned) #0
```


```{r}
#find num of missing vals
missing_counts <- colSums(is.na(cleaned))

#create df w/ col names and missing values 
missing_data <- data.frame(
  variable = names(missing_counts), 
  missing_values = missing_counts
)

print(missing_data)

```


Map values with more than 5% missing values 
```{r}
# Initialize an empty vector to store the names that satisfy the condition
selected_labels <- c()

# Loop through the list and add the names that meet the condition to the vector
for (label in names(missing_counts)) {
  value <- missing_counts[[label]]
  if (value > 4106) {
    selected_labels <- c(selected_labels, label)
  }
}

# Print the selected labels
print(selected_labels) #13
```

```{r}
map_data <- cleaned %>%
  select(x, y, all_of(selected_labels))

```


Notes: 
- NDVI var is not problematic; imputation is satisfactory 
- Two measurements taken across 10/18 at different times for most ECOSTRESS vars 
  - there is concerning amount/area of missingness for both of these times
- 10/22 has missingness more dispersed for most ECOSTRESS vars
  - 
```{r}
source("C:/Users/Bryant Willoughby/Documents/DRUMS/code/week 1/maps_function.R")

for (col_name in names(map_data)[3:ncol(map_data)]){
  #note: col_name is a character string of var names
  col <- map_data[[col_name]] #values of each variable 
  
  p <- map.heatmap(map_data$x, map_data$y, col, legendTitle = col_name, 
                   mainTitle = paste(col_name, "across region"))
  print(p)
}

```



Remove variables with most missingness (>2 missing)
- Imputing this number of missing values raises concern for biasing and incorrect analysis
```{r}
ncol(cleaned)#104

#remove variables with 'significant' missing data 
cleaned <- cleaned %>%
  select(-c("LST_2019.10.18.16.00.00", "LST_2019.10.18.23.00.00",  "ET_2019.10.18.16.00.00", "ET_2019.10.18.23.00.00", "ESI_2019.10.18.16.00.00", "ESI_2019.10.18.23.00.00", "WUE_2019.10.18.16.00.00", "WUE_2019.10.18.23.00.00"))

ncol(cleaned) #96

```
Check missingness again 
```{r}
#find num of missing vals
missing_counts <- colSums(is.na(cleaned))

#create df w/ col names and missing values 
missing_data <- data.frame(
  variable = names(missing_counts), 
  missing_values = missing_counts
)

```


```{r}
# Initialize an empty vector to store the names that satisfy the condition
selected_labels <- c()

# Loop through the list and add the names that meet the condition to the vector
for (label in names(missing_counts)) {
  value <- missing_counts[[label]]
  if (value > 4106) {
    selected_labels <- c(selected_labels, label)
  }
}

# Print the selected labels
print(selected_labels) #5
```

```{r}
map_data <- cleaned %>%
  select(x, y, all_of(selected_labels))

```


```{r}
source("C:/Users/Bryant Willoughby/Documents/DRUMS/code/week 1/maps_function.R")

for (col_name in names(map_data)[3:ncol(map_data)]){
  #note: col_name is a character string of var names
  col <- map_data[[col_name]] #values of each variable 
  
  p <- map.heatmap(map_data$x, map_data$y, col, legendTitle = col_name, 
                   mainTitle = paste(col_name, "across region"))
  print(p)
}

```

Perform KNN on dataset
```{r}
source("C:/Users/Bryant Willoughby/Documents/DRUMS/code/week 1/KNN_imputation_function.R")

for (i in 5:ncol(cleaned)){
  cleaned[[i]] <- knn(cleaned$x, cleaned$y, cleaned[[i]])
}

```

confirm there are no more missing values 
```{r}
#find num of missing vals
missing_counts <- colSums(is.na(cleaned))

#create df w/ col names and missing values 
missing_data <- data.frame(
  variable = names(missing_counts), 
  missing_values = missing_counts
)

print(missing_data)

anyNA(cleaned)

```

Rename x/y variables as long/lat respectively 
```{r}
cleaned <- cleaned %>%
  select(y, x, everything()) %>%
  rename("Lat" = y, "Lon" = x)

```


Write this cleaned data set to a csv file to work with
```{r}
#write.csv(cleaned, file = "C:/Users/Bryant Willoughby/Documents/DRUMS/raw(Kincade)/Final Data Sets/prefire_kincade_finaldata.csv", row.names = F)
          
```














