---
title: "Var_grouping"
author: "Bryant Willoughby"
date: "2023-06-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(readr)
library(ggplot2)
```

Say "original_variable" is observed at 100 locations at 5 times, indexed by lag so -7 means last week.  Then the intercepts and slopes at each site are (for fake data)
```{r}
 original_variable <- matrix(rnorm(500),100,5)  # Some fake data
 lag               <- c(-14,-10,-7,-4,-2)       # some fake time lags
  #specific time lag that is being considered as a predictor for response    var
 X                 <- matrix(0,100,2)           
 # These are the summaries we will use as features
 
 for(i in 1:100){
   X[i,] <- lm(original_variable[i,]~lag)$coef
 }

```

Then X[i,1] and X[i,2] are the current value and trend, respectively, for location i.  

Read in final data
```{r}
data <- read_csv("C:/Users/Bryant Willoughby/Documents/DRUMS/raw/Final data sets/prefire_knp_blr_finaldata.csv")

```

Process for grouping variables 
- Variables grouped across days: linear regression


Note: 9/9/2021 - lightning strike ignition
  - 31 days in Aug, 2021
  
Explore LST variable 
```{r}
LST <- data %>%
  select(starts_with("LST"))

names(LST)
```

```{r}
# Reshape the data to long format
LST_long <- LST %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

# Create a boxplot using ggplot2
ggplot(LST_long, aes(x = variable, y = value)) +
  geom_boxplot()

```


```{r}
firedate <- as.Date("2021-09-09")
dates <- as.Date(c("2021-08-26","2021-08-27","2021-08-30","2021-08-31",
                   "2021-09-03", "2021-09-04", "2021-09-07", "2021-09-08", 
                   "2021-09-09", "2021-09-13"))

diff <- dates - firedate
lag <- vector("numeric", length(dates))

for (i in 1:length(lag)){
  lag[i] <- diff[[i]]
}

original_variable <- as.matrix(LST)
X <- matrix(0,nrow(LST), 2)

for (i in 1:nrow(LST)){
  X[i,] <- lm(original_variable[i,] ~ lag)$coef
}


LST_sum <- X
colnames(LST_sum) <- c("LST_int", "LST_slope")
LST_sum <- data.frame(LST_sum)

```

ET variable
```{r}
ET <- data %>%
  select(starts_with("ET"))

names(ET)
```

```{r}
# Reshape the data to long format
ET_long <- ET %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

# Create a boxplot using ggplot2
ggplot(ET_long, aes(x = variable, y = value)) +
  geom_boxplot()

```


```{r}
firedate <- as.Date("2021-09-09")
dates <- as.Date(c("2021-08-26", "2021-08-30", "2021-09-03", "2021-09-07"))

diff <- dates - firedate

lag <- vector("numeric", length(dates))
for (i in 1:length(lag)){
  lag[i] <- diff[[i]]
}

original_variable <- as.matrix(ET)
X <- matrix(0,nrow(ET), 2)

for (i in 1:nrow(ET)){
  X[i,] <- lm(original_variable[i,] ~ lag)$coef
}

colnames(X) <- c("ET_int", "ET_slope")
ET_sum <- data.frame(X)

```

ESI variable 
```{r}
ESI <- data %>%
  select(starts_with("ESI"))

names(ESI)
```

```{r}
# Reshape the data to long format
ESI_long <- ESI %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

# Create a boxplot using ggplot2
ggplot(ESI_long, aes(x = variable, y = value)) +
  geom_boxplot()

```

```{r}
firedate <- as.Date("2021-09-09")
dates <- as.Date(c("2021-08-26", "2021-08-30", "2021-09-03", "2021-09-07"))

diff <- dates - firedate

lag <- vector("numeric", length(dates))
for (i in 1:length(lag)){
  lag[i] <- diff[[i]]
}

original_variable <- as.matrix(ESI)
X <- matrix(0,nrow(ESI), 2)

for (i in 1:nrow(ESI)){
  X[i,] <- lm(original_variable[i,] ~ lag)$coef
}

ESI_sum <- data.frame(X)
colnames(ESI_sum) <- c("ESI_int", "ESI_slope")

```

WUE
```{r}
WUE <- data %>%
  select(starts_with("WUE"))

names(WUE)
```

```{r}
# Reshape the data to long format
WUE_long <- WUE %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

# Create a boxplot using ggplot2
ggplot(WUE_long, aes(x = variable, y = value)) +
  geom_boxplot()

```

```{r}
firedate <- as.Date("2021-09-09")
dates <- as.Date(c("2021-08-26", "2021-08-30", "2021-09-03", "2021-09-07"))

diff <- dates - firedate

lag <- vector("numeric", length(dates))
for (i in 1:length(lag)){
  lag[i] <- diff[[i]]
}

original_variable <- as.matrix(WUE)
X <- matrix(0,nrow(WUE), 2)

for (i in 1:nrow(WUE)){
  X[i,] <- lm(original_variable[i,] ~ lag)$coef
}

WUE_sum <- data.frame(X)
colnames(WUE_sum) <- c("WUE_int", "WUE_slope")

```

LAI
- is this supposed to be 08/31/2021 or 1 second? 
- note: removing one variable I have questions abt for time being
```{r}
LAI <- data %>%
  select(starts_with("LAI"))

names(LAI)
```

```{r}
# Reshape the data to long format
LAI_long <- LAI %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

# Create a boxplot using ggplot2
ggplot(LAI_long, aes(x = variable, y = value)) +
  geom_boxplot()

```

```{r}
firedate <- as.Date("2021-09-09")
dates <- as.Date(c("2021-08-30", "2021-08-26", "2021-09-03", 
                   "2021-09-07", "2021-09-11"))

diff <- dates - firedate

lag <- vector("numeric", length(dates))
for (i in 1:length(lag)){
  lag[i] <- diff[[i]]
}

original_variable <- as.matrix(LAI)
X <- matrix(0,nrow(LAI), 2)

for (i in 1:nrow(LAI)){
  X[i,] <- lm(original_variable[i,] ~ lag)$coef
}

LAI_sum <- data.frame(X)
colnames(LAI_sum) <- c("LAI_int", "LAI_slope")

```


NDVI
```{r}
NDVI <- data %>%
  select(starts_with("NDVI"))

names(NDVI)
```

```{r}
# Reshape the data to long format
NDVI_long <- NDVI %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

# Create a boxplot using ggplot2
ggplot(NDVI_long, aes(x = variable, y = value)) +
  geom_boxplot()

```
```{r}
firedate <- as.Date("2021-09-09")
dates <- as.Date(c("2021-08-30", "2021-08-31", "2021-09-04",
                   "2021-09-09"))

diff <- dates - firedate

lag <- vector("numeric", length(dates))
for (i in 1:length(lag)){
  lag[i] <- diff[[i]]
}

original_variable <- as.matrix(NDVI)
X <- matrix(0,nrow(NDVI), 2)

for (i in 1:nrow(NDVI)){
  X[i,] <- lm(original_variable[i,] ~ lag)$coef
}

NDVI_sum <- data.frame(X)
colnames(NDVI_sum) <- c("NDVI_int", "NDVI_slope")

```

AM_soil_moisture
```{r}
AM_sm <- data %>%
  select(starts_with("AM"))

names(AM_sm)
```

```{r}
# Reshape the data to long format
AM_sm_long <- AM_sm %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

# Create a boxplot using ggplot2
ggplot(AM_sm_long, aes(x = variable, y = value)) +
  geom_boxplot()

```

```{r}
firedate <- as.Date("2021-09-09")
dates <- as.Date(c("2021-08-29", "2021-09-03", "2021-09-05",
                   "2021-09-06", "2021-09-08", "2021-09-11"))

diff <- dates - firedate

lag <- vector("numeric", length(dates))
for (i in 1:length(lag)){
  lag[i] <- diff[[i]]
}

original_variable <- as.matrix(AM_sm)
X <- matrix(0,nrow(AM_sm), 2)

for (i in 1:nrow(AM_sm)){
  X[i,] <- lm(original_variable[i,] ~ lag)$coef
}

AM_sm_sum <- data.frame(X)
colnames(AM_sm_sum) <- c("AM_sm_int", "AM_sm_slope")

```

PM soil moisture 
```{r}
PM_sm <- data %>%
  select(starts_with("soil"))

names(PM_sm)
```

```{r}
# Reshape the data to long format
PM_sm_long <- PM_sm %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

# Create a boxplot using ggplot2
ggplot(PM_sm_long, aes(x = variable, y = value)) +
  geom_boxplot()

```

```{r}
firedate <- as.Date("2021-09-09")
dates <- as.Date(c("2021-08-30", "2021-09-01", "2021-09-02",
                   "2021-09-04", "2021-09-07", "2021-09-09", 
                   "2021-09-10"))

diff <- dates - firedate

lag <- vector("numeric", length(dates))
for (i in 1:length(lag)){
  lag[i] <- diff[[i]]
}

original_variable <- as.matrix(PM_sm)
X <- matrix(0,nrow(PM_sm), 2)

for (i in 1:nrow(PM_sm)){
  X[i,] <- lm(original_variable[i,] ~ lag)$coef
}

PM_sm_sum <- data.frame(X)
colnames(PM_sm_sum) <- c("PM_sm_int", "PM_sm_slope")

```

Dewpoint temp (tdmean)
```{r}
tdmean <- data %>%
  select(starts_with("tdmean"))

names(tdmean)
```

```{r}
# Reshape the data to long format
tdmean_long <- tdmean %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

# Create a boxplot using ggplot2
ggplot(tdmean_long, aes(x = variable, y = value)) +
  geom_boxplot()

```

```{r}
firedate <- as.Date("2021-09-09")
dates <- as.Date(c("2021-08-29", "2021-08-30", "2021-08-31",
                   "2021-09-01", "2021-09-02", "2021-09-03", 
                   "2021-09-04", "2021-09-05", "2021-09-06", 
                   "2021-09-07", "2021-09-08", "2021-09-09", 
                   "2021-09-10", "2021-09-11", "2021-09-12"))

diff <- dates - firedate

lag <- vector("numeric", length(dates))
for (i in 1:length(lag)){
  lag[i] <- diff[[i]]
}

original_variable <- as.matrix(tdmean)
X <- matrix(0,nrow(tdmean), 2)

for (i in 1:nrow(tdmean)){
  X[i,] <- lm(original_variable[i,] ~ lag)$coef
}

tdmean_sum <- data.frame(X)
colnames(tdmean_sum) <- c("tdmean_int", "tdmean_slope")

```

Air temp (tmean)
```{r}
tmean <- data %>%
  select(starts_with("tmean"))

names(tmean)
```

```{r}
# Reshape the data to long format
tmean_long <- tmean %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

# Create a boxplot using ggplot2
ggplot(tmean_long, aes(x = variable, y = value)) +
  geom_boxplot()

```


```{r}
firedate <- as.Date("2021-09-09")
dates <- as.Date(c("2021-08-29", "2021-08-30", "2021-08-31",
                   "2021-09-01", "2021-09-02", "2021-09-03", 
                   "2021-09-04", "2021-09-05", "2021-09-06", 
                   "2021-09-07", "2021-09-08", "2021-09-09", 
                   "2021-09-10", "2021-09-11", "2021-09-12"))

diff <- dates - firedate

lag <- vector("numeric", length(dates))
for (i in 1:length(lag)){
  lag[i] <- diff[[i]]
}

original_variable <- as.matrix(tmean)
X <- matrix(0,nrow(tmean), 2)

for (i in 1:nrow(tmean)){
  X[i,] <- lm(original_variable[i,] ~ lag)$coef
}

tmean_sum <- data.frame(X)
colnames(tmean_sum) <- c("tmean_int", "tmean_slope")

```

Vapor Pressure Max (vpdmax)
```{r}
vpdmax <- data %>%
  select(starts_with("vpdmax"))

names(vpdmax)
```

```{r}
# Reshape the data to long format
vpdmax_long <- vpdmax %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

# Create a boxplot using ggplot2
ggplot(vpdmax_long, aes(x = variable, y = value)) +
  geom_boxplot()

```

```{r}
firedate <- as.Date("2021-09-09")
dates <- as.Date(c("2021-08-29", "2021-08-30", "2021-08-31",
                   "2021-09-01", "2021-09-02", "2021-09-03", 
                   "2021-09-04", "2021-09-05", "2021-09-06", 
                   "2021-09-07", "2021-09-08", "2021-09-09", 
                   "2021-09-10", "2021-09-11", "2021-09-12"))

diff <- dates - firedate

lag <- vector("numeric", length(dates))
for (i in 1:length(lag)){
  lag[i] <- diff[[i]]
}

original_variable <- as.matrix(vpdmax)
X <- matrix(0,nrow(vpdmax), 2)

for (i in 1:nrow(vpdmax)){
  X[i,] <- lm(original_variable[i,] ~ lag)$coef
}

vpdmax_sum <- data.frame(X)
colnames(vpdmax_sum) <- c("vpdmax_int", "vpdmax_slope")

```

Vapor Pressure Min (vpdmin)
```{r}
vpdmin <- data %>%
  select(starts_with("vpdmin"))

names(vpdmin)
```

```{r}
# Reshape the data to long format
vpdmin_long <- vpdmin %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

# Create a boxplot using ggplot2
ggplot(vpdmin_long, aes(x = variable, y = value)) +
  geom_boxplot()

```

```{r}
firedate <- as.Date("2021-09-09")
dates <- as.Date(c("2021-08-29", "2021-08-30", "2021-08-31",
                   "2021-09-01", "2021-09-02", "2021-09-03", 
                   "2021-09-04", "2021-09-05", "2021-09-06", 
                   "2021-09-07", "2021-09-08", "2021-09-09", 
                   "2021-09-10", "2021-09-11", "2021-09-12"))

diff <- dates - firedate

lag <- vector("numeric", length(dates))
for (i in 1:length(lag)){
  lag[i] <- diff[[i]]
}

original_variable <- as.matrix(vpdmin)
X <- matrix(0,nrow(vpdmin), 2)

for (i in 1:nrow(vpdmin)){
  X[i,] <- lm(original_variable[i,] ~ lag)$coef
}

vpdmin_sum <- data.frame(X)
colnames(vpdmin_sum) <- c("vpdmin_int", "vpdmin_slope")

```


Create df w/ slope & intercept term
```{r}
newdata <- data %>%
  select("Lat", "Lon", "dnbr6", "dnbr", 
         "aspect", "slope", "elev", "LC")

fulldata <- cbind(newdata, LST_sum, ET_sum, ESI_sum, 
                  WUE_sum, LAI_sum, NDVI_sum, AM_sm_sum, 
                  PM_sm_sum, tdmean_sum, tmean_sum, vpdmax_sum, 
                  vpdmin_sum)

```

Write to CSV to experiment with 
```{r}
#write.csv(fulldata, "C:/Users/Bryant Willoughby/Documents/DRUMS/raw/intermediate/prefire_knp_blr_grouped_finaldata.csv", row.names = F)
```
