---
title: "Windy_EDA"
author: "Bryant Willoughby"
date: "2023-07-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
```

Read in Data 

```{r}
data <- read_csv("C:/Users/Bryant Willoughby/Documents/DRUMS/raw(Windy)/Final Data Sets/prefire_windy_fulldata.csv")
data

```

Check to see if there are any completely missing rows (records)
- If so, remove these up front 
- note: there were 168 records with completley missing data that I removed 
```{r}
head(data)

#confirm no missing vals in first 4 vars
test <- complete.cases(data[,(1:4)])
anyNA(test) #FALSE

#keeping rows that have complete information in all columns (except the first four) or have at least one non-missing value in any of the columns (excluding the first four columns).
cleaned <- data[complete.cases(data[,-(1:4)]) | rowSums(!is.na(data[,-(1:4)])) > 0, ]

# Print the updated data frame
head(cleaned)

nrow(data) - nrow(cleaned) #168
```

```{r}
#find num of missing vals
missing_counts <- colSums(is.na(cleaned))

#create df w/ col names and missing values 
missing_data <- data.frame(
  variable = names(missing_counts), 
  missing_values = missing_counts
)

print(missing_data)

```

```{r}
# Initialize an empty vector to store the names that satisfy the condition
selected_labels <- c()

# Loop through the list and add the names that meet the condition to the vector
for (label in names(missing_counts)) {
  value <- missing_counts[[label]]
  if (value > 3000) {
    selected_labels <- c(selected_labels, label)
  }
}

# Print the selected labels
print(selected_labels)
```

```{r}
map_data <- data %>%
  select(x, y, all_of(selected_labels))

```



```{r}
source("C:/Users/Bryant Willoughby/Documents/DRUMS/code/week 1/maps_function.R")

for (col_name in names(map_data)[3:6]){
  #note: col_name is a character string of var names
  col <- map_data[[col_name]] #values of each variable 
  
  p <- map.heatmap(map_data$x, map_data$y, col, legendTitle = col_name, 
                   mainTitle = paste(col_name, "across region"))
  print(p)
}

```

Remove variables with most missingness (>48,000 missing)
- Imputing this number of missing values raises concern for biasing and incorrect analysis
  - LST_2021.08.28.10.00.00 (77371)
  - LST_2021.09.08.07.00.00 (97914)
```{r}
ncol(cleaned) #106

#remove variables with 'significant' missing data 
cleaned <- cleaned %>%
  select(-c(LST_2021.08.28.10.00.00, LST_2021.09.08.07.00.00,))

ncol(cleaned) #104

```
remind myself what is still missing 
- notice: there is no more than 4,274 missing obs from any variable 
- allows for better KNN procedure to use local data point estimates for imputation 
- now replacing this missing data w/ an appropriately estimated value 
```{r}
#find num of missing vals
missing_counts <- colSums(is.na(cleaned))

#create df w/ col names and missing values 
missing_data <- data.frame(
  variable = names(missing_counts), 
  missing_values = missing_counts
)

print(missing_data)

```

Now conduct KNN imputation on all of the missing data points 

KNN: 
1. set K to choose number of neighbors
2. calculate distances between target (missing) point and all other points in dataset
3. sort the calculated distances
4. getting the labels of the top K entries 
5. return prediction about the target (missing) point
```{r}
#load function to use for imputation 
knn <- function(long,lat,X,k=10){
  miss <- which(is.na(X))
  full <- X
  for(i in miss){
    d       <- sqrt((long[i] - long)^2 + (lat[i]  - lat)^2)
    d[miss] <- Inf
    neigh   <- order(d)[1:k]
    full[i] <- mean(X[neigh])
  }
  return(full)
  }
```

Perform KNN on dataset
```{r}
for (i in 5:ncol(cleaned)){
  cleaned[[i]] <- knn(cleaned$x, cleaned$y, cleaned[[i]])
}

```

confirm there are no more missing values 
```{r}
#find num of missing vals
missing_counts <- colSums(is.na(cleaned))

#create df w/ col names and missing values 
missing_data <- data.frame(
  variable = names(missing_counts), 
  missing_values = missing_counts
)

print(missing_data)

anyNA(cleaned)

```


Rename x/y variables as long/lat respectively 
```{r}
cleaned <- cleaned %>%
  select(y, x, everything()) %>%
  rename("Lat" = y, "Lon" = x)

```

Write this cleaned data set to a csv file to work with
```{r}
#write.csv(cleaned, file = "C:/Users/Bryant Willoughby/Documents/DRUMS/raw(Windy)/Final Data Sets/prefire_windy_finaldata.csv", row.names = F)
          
```










